Measuring neutron star tidal deformability with Advanced LIGO: a Bayesian analysis of neutron star - black hole binary observations

### 2016-09-05 Paper/paper.tex
improved s3 more

Next, in Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the ratio of
measurement errors $\arr_\eta$ for the symmetric mass-ratio.
% 
Going through the figure from left to right, we find that for realistic
SNRs ($\rho\leq 30$) the systematics remain below statistical errors for
$\eta$ measurement. The worst case is of the most deformable NSs
($\lambdans = 2000$), but even for them systematics in $\eta$ are $2\times$
smaller than the statistical measurement errors.
% 
Moving to louder signals with $\rho\simeq 50$, we find that for binaries of
fairly deformable NSs ($\lambdans\gtrsim 1500$) and low-mass BHs
($\mbh\leq 5M_\odot$) that have prograde spins ($\chibh\gtrsim +0.4$), our
measurement of mass-ratio can be seriously compromised by ignoring tidal
physics in template models. 
% 
This pattern is continued at even higher SNRs, as we can see from
Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. We therefore conclude that, even if
under moderate restrictions on BH and NS parameters, $\rho=50$ is loud enough
to motivate the use of tidal templates in aLIGO data analyses.
% 
In addition, we also notice that, unlike for $\mchirp$, the measured value of
$\eta$ is always {\it lower} than its true value, which is what we expect if we
want BBH templates to fit NSBHs that disrupt and merge at lower frequencies.

Moving on from mass to spin parameters, we now consider the measurement of BH
spin angular momentum $\chibh$. The ratio of systematic and statistical errors
associated with $\chibh$ measurement is shown in
Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR}. The presentation of information in this 
figure is identical to that of Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. A diverging colormap is used so that both 
extremes of the colorbar range point to large systematic biases, while its zero (or
small) value lies in the middle.
% 
For the lowest SNR considered ($\rho=30$), $\chibh$ bias is about $2\times$
smaller than its statistical measurement uncertainty, and is therefore mostly
negligible. Both do become somewhat comparable, but only when we have the most
deformable NSs in orbit around low-mass BHs. 
% 
At higher SNRs $(\rho\simeq50-70)$, we find that the systematics in $\chibh$
measurement can dominate completely, especially for binaries containing
mass-gap violating BHs and/or deformable NSs with $\lambdans\geq1000$.
% 
From Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR} we additionally note that when
the source spin magnitudes approach the highest allowed, i.e. at both extremes
of the $x$-axes, $\chibh\times\arr_{\chibh}<0$. This is to be expected because
the median of the recovered posterior distributions for $\chibh$ can only get
pushed inwards from the boundaries.

Summarizing these results, we find that irrespective of system parameters, below a
signal-to-noise ratio of $30$, our measurements of mass and spin parameters of 
astrophysical NSBH binaries will remain limited by the intrinsic uncertainty due to 
instrument noise, and do not depend on whether we include tidal effects in template
models. However, when the signal-to-noise ratio exceeds $30$ the systematic bias in
binary mass and spin measurements become comparable to and can exceed the
uncertainty due to noise. Of the different parameters considered, we find that
amongst all non-tidal parameters, the measurement of $\eta$ degrades worst (in a 
relative-error sense) due to the use of BBH templates in deciphering an NSBH
signal.

>>>>

Next, in Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}, we show the ratio of
measurement errors $\arr_\eta$ for the symmetric mass-ratio.
% 
Going through the figure from left to right, we find that for realistic
SNRs ($\rho\leq 30$) the systematics remain below statistical errors for
$\eta$ measurement. The worst case is of the most deformable NSs
($\lambdans = 2000$), but even for them systematics in $\eta$ are $2\times$
smaller than the statistical measurement errors.
% 
Moving to louder signals with $\rho\simeq 50$, we find that for binaries of
fairly deformable NSs ($\lambdans\gtrsim 1500$) and low-mass BHs
($\mbh\leq 5M_\odot$) that have prograde spins ($\chibh\gtrsim +0.4$), our
measurement of mass-ratio can be seriously compromised by ignoring tidal
physics in template models. 
% 
This pattern is continued at even higher SNRs, as we can see from
Fig.~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. We therefore conclude that, even if
under moderate restrictions on BH and NS parameters, $\rho=30-50$ is loud enough
to motivate the use of tidal templates in aLIGO data analyses.
% 
In addition, we also notice that, unlike for $\mchirp$, the measured value of
$\eta$ is always {\it lower} than its true value, which is what we expect if we
want BBH templates to fit NSBHs that disrupt and merge at lower frequencies.

Moving on from mass to spin parameters, we now consider the measurement of BH
spin angular momentum $\chibh$. The ratio of systematic and statistical errors
for $\chibh$ are shown in
Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR}. The presentation of information in this 
figure is identical to that of Fig.~\ref{fig:TN_chirpMassBias_vs_Lambda_SNR}
and~\ref{fig:TN_EtaBias_vs_Lambda_SNR}. A diverging colormap is used so that both 
extremes of the colorbar range point to large systematic biases, while its zero (or
small) value lies in the middle.
% 
For the lowest SNR considered ($\rho=30$), $\chibh$ bias is about $2\times$
smaller than its statistical measurement uncertainty, and is therefore mostly
negligible. Both do become somewhat comparable, but only when we have the most
deformable NSs in orbit around low-mass BHs. 
% 
At higher SNRs $(\rho\simeq50-70)$, we find that the systematics in $\chibh$
measurement can dominate completely, especially for binaries containing
mass-gap violating BHs and/or deformable NSs with $\lambdans\geq1000$.
% 
From Fig.~\ref{fig:TN_BHspinBias_vs_Lambda_SNR} we additionally note that when
the source spin magnitudes approach the highest allowed, i.e. at both extremes
of the $x$-axes, $\chibh\times\arr_{\chibh}<0$. This is to be expected because
the median of the recovered posterior distributions for $\chibh$ can only get
pushed inwards from the boundaries.

Summarizing these results, we find that irrespective of system parameters,
below a signal-to-noise ratio of $30$, our measurements of mass and spin
parameters of astrophysical NSBH binaries will remain limited by the intrinsic
uncertainty due to instrument noise, and do not depend on whether we include
tidal effects in template models. However, when the signal-to-noise ratio
exceeds $30$ the systematic bias in binary mass and spin measurements become
comparable to and can exceed the uncertainty due to noise. Of the different
non-tidal parameters considered, we find that the measurement of $\eta$
degrades worst (in a relative-error sense) due to the use of BBH templates in
deciphering an NSBH signal. Of all the sub-categories, we find that tidal
templates could especially help with the parameter estimation of astrophysical
{\it mass-gap violating} NSBH binaries,
